---
title: "Integration_Tumor"
author: "Camino RSM"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

## Load Seurat, add metadata, remove doublets
```{r}
library(Seurat)
library(harmony)
library(SeuratDisk)
library(dplyr)
Tumor = readRDS("/home/caminors/Desktop/sc_RNAseq/QC_seurats/SARC_Tumor_f_Soup_doublet.rds")
Tumor@assays$SCT = NULL
DefaultAssay(Tumor)="RNA"
metadata = read.csv("/home/caminors/Desktop/sc_RNAseq/QC_seurats/PDSC_scRNAseq_metadata_VY20241003.csv")
new_metadata <- metadata %>%
  distinct(biosample, .keep_all = TRUE) %>%  # Keep only unique `biosample` rows
  mutate(
    biosample = case_when(
      biosample == "hSC26_adj" ~ "hSC26_AdjNorm",  # Replace specific value
      biosample == "hSC26_t" ~ "hSC26_ddLPS",     # Replace another specific value
      TRUE ~ biosample                             # Keep the rest unchanged
    )
  )
new_metadata$Run = NULL
rownames(new_metadata)=new_metadata$biosample
new_metadata$orig.ident = new_metadata$biosample

updated_metadata <- merge(
  x = Tumor@meta.data, 
  y = new_metadata, 
  by = "orig.ident", 
  all.x = TRUE  # Keeps all cells in the Seurat object
)

# Restore row names (critical for Seurat objects)
rownames(updated_metadata) <- rownames(Tumor@meta.data)
# Update the Seurat object's metadata
Tumor@meta.data <- updated_metadata
table(Tumor$orig.ident)

Tumor = subset(Tumor, scDblFinder.class %in% c("singlet"))

table(Tumor$orig.ident)
```

### Remove ENSG genes that were unmapped from all assays
```{r}
# Extract gene names starting with "ENSG" from the Tumor object
ENS_genes <- Tumor@assays[["RNA"]]@counts@Dimnames[[1]][grepl("^ENSG", Tumor@assays[["RNA"]]@counts@Dimnames[[1]])]
# Extract gene names from SCT assay
Total_genes <- rownames(Tumor@assays[["RNA"]]@counts)
# Remove the identified genes from the SCT assay
Tumor@assays[["RNA"]] <- subset(Tumor@assays[["RNA"]], features = Total_genes[!Total_genes %in% ENS_genes])
cat("Number of genes removed:", (length(Total_genes) -nrow(Tumor@assays[["RNA"]])), "\n")
```

## Cell-type prediction using Azimuth
```{r}
library(Azimuth)
DefaultAssay(Tumor)="RNA"
Tumor <- RunAzimuth(Tumor, reference="adiposeref")
table(Tumor$predicted.celltype.l1)
```
## Splitting the RNA assay on layers by orig.ident --> one count and one data for each orig.ident
```{r}
Tumor
Tumor[["RNA"]] <- split(Tumor[["RNA"]], f = Tumor$orig.ident)
Tumor
```
# Normalize Data using SCT
```{r}
library(future)
options(future.globals.maxSize=30000*1024^2)
Tumor <- SCTransform(Tumor, assay="RNA")
```

## Remove genes known to be distorting the clustering (HSP/VDJ genes)
```{r}
### Remove genes that distort the clustering (HSP/VDJ genes)
vdj_genes <- c(
    "TRAV1-1", "TRAV1-2", "TRAV1-3", "TRAV2", "TRAV3", "TRAV4", "TRAV5", "TRAV6", "TRAV7", "TRAV8",
    "TRBV1", "TRBV2", "TRBV3", "TRBV4", "TRBV5", "TRBV6", "TRBV7", "TRBV8", "TRBV9", "TRBV10",
    "TRGV1", "TRGV2", "TRGV3", "TRGV4", "TRGV5", "TRGV6", "TRGV7", "TRGV8",
    "TRDV1", "TRDV2", "TRDV3", "TRDV4",
    "IGHV1-1", "IGHV1-2", "IGHV1-3", "IGHV2-5", "IGHV2-6", "IGHV3-1", "IGHV3-2", "IGHV3-3",
    "IGKV1-1", "IGKV1-2", "IGKV1-3", "IGKV2-5", "IGKV2-6", "IGKV3-1", "IGKV3-2", "IGKV3-3",
    "IGLV1-1", "IGLV1-2", "IGLV1-3", "IGLV2-14", "IGLV2-15", "IGLV3-1", "IGLV3-2", "IGLV3-3"
)
hsp_genes <- c(
    "HSPA1A", "HSPA1B", "HSPA2", "HSPA5", "HSPA6", "HSPA7", "HSPA8", "HSPA9", "HSPA12A", "HSPA12B",
    "HSP90AA1", "HSP90AB1", "HSP90B1", "HSP90C1",
    "HSPD1", "CCT2", "CCT3", "CCT4", "CCT5", "CCT6A", "CCT6B",
    "DNAJA1", "DNAJA2", "DNAJA3", "DNAJB1", "DNAJB2", "DNAJB4", "DNAJB5", "DNAJC1", "DNAJC2",
    "HSPE1", 
    "HSPB1", "HSPB2", "HSPB3", "HSPB6", "HSPB7", "HSPB8", "HSPB9", "HSPB11",
    "HSPH1", "HSPH2", "HSPC1", "HSPC2"
)


# Combine all genes to be removed (VDJ, HSP)
genes_to_remove <- unique(c(vdj_genes, hsp_genes))

# get genes present in SCT
SCT_genes =Tumor@assays[["SCT"]]@var.features
# Identify which of the genes to be removed are present in the SCT assay
genes_in_SCT <- SCT_genes[SCT_genes %in% genes_to_remove]

# Count how many genes are being removed
num_genes_removed <- length(genes_in_SCT)

# Remove the identified genes 
Tumor@assays[["SCT"]]@var.features <- SCT_genes[!SCT_genes %in% genes_in_SCT]

# Print the number of genes removed
cat("Number of genes removed from the SCT assay:", num_genes_removed, "\n")
```

```{r}
Tumor <- RunPCA(Tumor)
Tumor <- IntegrateLayers(
  object = Tumor, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "harmony",
  verbose = FALSE
)
```
## Check Elbow Plot
```{r}
ElbowPlot(Tumor, ndims = 50)
```
```{r, fig.width=20, fig.height=5}
Tumor <- FindNeighbors(Tumor, reduction = "harmony", dims = 1:20)
Tumor <- FindClusters(Tumor, resolution = 0.5, cluster.name = "harmony_cluster")
Tumor <- RunUMAP(Tumor, reduction = "harmony", dims = 1:20, reduction.name = "umap.harmony")
DimPlot(Tumor,
  reduction = "umap.harmony",
  group.by = c("orig.ident", "predicted.celltype.l1", "harmony_cluster"),
  combine = TRUE, label.size = 2
)
```



## Mnn Integration
```{r, fig.width=20, fig.height=5}
library(SeuratWrappers)
DefaultAssay(Tumor)="SCT"
Tumor <- IntegrateLayers(
  object = Tumor, method = FastMNNIntegration,
  new.reduction = "Mnn",
  verbose = FALSE,  batch= Tumor$orig.ident )
Tumor <- FindNeighbors(Tumor, reduction = "Mnn", dims = 1:20)
Tumor <- FindClusters(Tumor, resolution = 0.5, cluster.name = "mnn_cluster")
Tumor <- RunUMAP(Tumor, reduction = "Mnn", dims = 1:20, reduction.name = "umap.mnn")

DimPlot(Tumor,
  reduction = "umap.mnn",
  group.by = c("orig.ident", "predicted.celltype.l1", "mnn_cluster"),
  combine = TRUE, label.size = 2
)
```


## Calculate lisi score
```{r,  fig.width=6, fig.height=7}

Tumor$lisi_harmony <- lisi::compute_lisi(Tumor@reductions[["umap.harmony"]]@cell.embeddings, Tumor@meta.data, "orig.ident")
Tumor$lisi_mnn <- lisi::compute_lisi(Tumor@reductions[["umap.mnn"]]@cell.embeddings, Tumor@meta.data, "orig.ident")

## Higher lisi means better mix between samples :)
summary(Tumor$lisi_harmony)
summary(Tumor$lisi_mnn)
library(ggplot2)
# Generate the plots
FeaturePlot(Tumor, features = c("lisi_harmony"), reduction = "umap.harmony") + FeaturePlot(Tumor, features=c("lisi_mnn"), reduction="umap.mnn")
```
## Save object
```{r}
saveRDS(Tumor, "./Tumor_umap_harmony_mnn_Azimuth.rds")
```

